% !TEX TS-program = xelatex
% File: house-style-and-preamble.tex
% Purpose: Brett's typical preamble + usage conventions:
%   examples (langsci-gb4e), italics vs. quotation, dashes, contractions, plain style, etc.

\documentclass[12pt]{article}

% == Layout & language ==
\usepackage[margin=1in]{geometry}
\usepackage[british]{babel}            % British conventions; use -ize spellings in prose
\usepackage{fontspec}                  % Xe/LuaLaTeX
\setmainfont{Charis SIL}
\usepackage[final]{microtype}

% == Quotation marks ==
\usepackage{csquotes}                  % \enquote{\ldots{}} with locale-aware quoting

% == Hyperlinks ==
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue,
  pdfauthor={Brett Reynolds},
  pdftitle={House Style \& Preamble}
}

% == Numbered linguistic examples (LangSci/gb4e wrapper, no 'exe' env) ==
\usepackage{langsci-gb4e}
\noautomath

% == Lists & small utilities ==
\usepackage{enumitem}
\setlist{itemsep=0.3\baselineskip, topsep=0.3\baselineskip}
\usepackage{xspace}

% =========================
% Bibliography (biblatex)
% =========================
% Default portable setup:
\usepackage[backend=biber,style=apa,natbib=true,doi=true,isbn=false,url=true]{biblatex}
\addbibresource{references.bib}

% If working in LangSci projects, you can switch to their unified style:
% \usepackage[backend=biber,style=unified,natbib=true,doi=true,isbn=false,url=false]{biblatex}

% =========================
% Light house macros
% =========================
% Mention (italics) vs. quotation marks
\newcommand{\term}[1]{\textit{#1}}

% Small-caps abbreviations for glosses
\newcommand{\abbr}[1]{\textsc{#1}}

% Judgement markers
\newcommand{\ungram}[1]{*\!#1}
\newcommand{\marg}[1]{?\!#1}
\newcommand{\odd}[1]{\#\!#1}

% e.g., i.e., etc., with sensible spacing
\newcommand{\eg}{e.g.,\xspace}
\newcommand{\ie}{i.e.,\xspace}
\newcommand{\etc}{etc.\xspace}

% =========================
% Document
% =========================
\title{House Style \& Preamble:\\A Compact Demonstration}
\author{Brett Reynolds}
\date{\today}

\begin{document}
\maketitle

\section{Mention vs.\ quotation}
Use italics for mention of forms and categories; reserve quotation marks for actual quotations.
Mention: \term{going forward} is best analysed as a preposition selecting a directional PP.
Quotation (with \verb|\enquote|): As \enquote{going forward} spread in business English, its syntax broadened.
Nested quotes: \enquote{Outer with an \enquote{inner} quote}.

\section{Dashes, ranges, hyphens}
Use an en dash with a nonbreaking space before it: the category is stable~-- within limits~-- across registers.
Use an en dash for ranges and relations: 2001--2025; pp.\ 113--127; the form--function distinction.
Use hyphens only for compounds: corpus-based study, task-specific rubric.

\section{Contractions and plain style}
Contractions are fine where rhythm improves: \enquote{We don't gain anything by multiplying categories.}
Prefer direct verbs and short clauses. Delete throat-clearers:
\begin{itemize}[nosep]
  \item \textbf{Avoid:} \enquote{It is important to note that the results clearly show that \dots}
  \item \textbf{Prefer:} \enquote{The results show \dots}
\end{itemize}

\section{Numbered examples with \texttt{langsci-gb4e}}
Single example (note: no \texttt{exe} environment):
\ea\label{ex:simple}
The committee have decided to adjourn.
\z

Judgement markers and subexamples:
\ea\label{ex:sub}
\ex She has already left.
\ex[*]{\ungram{She have already left.}}
\ex[\#]{\odd{The square triangle laughed.}}
\z

Cross-reference: see (\ref{ex:sub}).

Interlinear gloss (\texttt{\textbackslash gll}/\texttt{\textbackslash glt}; abbreviations in small caps):
\ea\label{ex:gloss}
\gll Ich sehe den Hund. \\
     I see  the.\abbr{acc} dog \\
\glt `I see the dog.'
\z

\section{Category vs.\ function}
Distinguish lexical categories (e.g., \term{determinative}, \term{adjective}) from functions (e.g., \term{head}, \term{modifier}, \term{subject}).
In \enquote{\term{those} books}, \term{those} is a \term{determinative} (category) functioning as a \term{determinative} (function) of the NP.

\section{Minimal citation examples (biblatex)}
Parenthetical: \citep{HuddlestonPullum2002}. Textual: \textcite{Pullum2018}.
Multiple: \citep{HuddlestonPullum2002,Pullum2018}. Page: \citep[113--127]{HuddlestonPullum2002}.

% \printbibliography % enable when you want the list

\end{document}
